---
title: "Open Source datasets"
author: "Pilii"
date: "6/21/2021"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: cerulean
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(USF)
library(readxl)
library(kableExtra)
#usethis::use_data_raw(name = 'name')
```

The following is a descriptive document of the open source datasets that might be of use for the USF summer institute.

# <ins> On Police shootings/killings </ins>

### Washington Post Data - police shootings
Source https://github.com/washingtonpost/data-police-shootings

```{r echo=FALSE}
data(wpdata)
print(colnames(wpdata))
```

### Fatal Encounters
Source https://fatalencounters.org/spreadsheets/

```{r echo=FALSE}
data(encounters)
colnames(encounters)
```

### Million Dollar Hoods
Source: https://milliondollarhoods.pre.ss.ucla.edu/test-iframe-2/

LAPD 2012-2017 by Neighborhoods (Ranking)
```{r echo=FALSE}
data(mdh_lapd)
colnames(mdh_lapd)
```

LASD 2012-2017 by Neighborhoods (Ranking)
```{r echo=FALSE}
data(mdh_lasd)
colnames(mdh_lasd)
```

### Mapping Police Violence
Source: https://mappingpoliceviolence.org/aboutthedata

Whole data set
```{r}
 data(mpv)
```

Worksheet 1 "2013-2020 Police Killings"
```{r echo=FALSE}
data(mpv_polkill)
colnames(mpv_polkill)
```

Worksheet 2 "2013-2020 Killings by PD"
```{r echo=FALSE}
data(mpv_killbypd)
colnames(mpv_killbypd)
```

Worksheet 3 "2013-2020 Killings by State"
```{r echo=FALSE}
data(mpv_killbystate)
colnames(mpv_killbystate)
```

Worksheet 4 "Police Killings of Black Men"
```{r echo=FALSE}
data(mpv_polkilbm)
colnames(mpv_polkilbm)
```


### Police 


# <ins> Policing in general </ins>


### Timeline
http://criticalresistance.org/policing-timeline/


### Spending

How much Us cities spend on policing
https://www.statista.com/chart/10593/how-much-do-us-cities-spend-on-policing/

PopularDemocracy.org
https://www.populardemocracy.org/news-and-publications/congress-must-divest-billion-dollar-police-budget-and-invest-public-education


### Stops - Open Policing Stanford
Source: https://openpolicing.stanford.edu/data/
CSV and shapefiles for select locations.


Los Angeles example
```{r echo=FALSE}
data(laops)
colnames(laops)
```


### Police in schools


Policing America’s Schools: An Education Week Analysis
U.S. and States
Source: https://www.edweek.org/which-students-are-arrested-most-in-school-u-s-data-by-school#/overview

### Jails and Prisons


### National Neighbourhood Watch
https://nnw.org/find-a-watch-program?type=2&field_county_value=&field_address_locality=&field_address_postal_code=&field_address_administrative_area=


### Critical Resistance - Abolishing Policing
Source: http://criticalresistance.org/abolish-policing/

#### On the Road with Abolition - ppt https://docs.google.com/presentation/d/1Vs1Dt53DR6s2vShPZw5xNX-QCDC2Ppws07ctl6h_Weg/edit#slide=id.g80c2475b0e_1_80

"the most basic police interaction, the stop"


# <ins> Drawing the Black Box </ins>

## <ins> Property data </ins>

Most cities have property data available you can download for free: "Property Vaulation and Assessment Data". 

### NYC Real Estate Assessment Property data.
Built with the Department of Finance values.
Source: https://data.cityofnewyork.us/City-Government/Property-Valuation-and-Assessment-Data/yjxr-fw8i

### MassGIS - MA Interactive Property Map
Datasets per city ready for download.
Source: https://massgis.maps.arcgis.com/apps/OnePane/basicviewer/index.html?appid=47689963e7bb4007961676ad9fc56ae9

### LA County Open Data Initiative
Property related data: https://data.lacounty.gov/browse?category=Property%2FPlanning&utf8=%E2%9C%93

### Zillow
Source: https://www.zillow.com/research/data/

"Zillow publishes top-tier ZHVI (typical value for homes within the 65th to 95th percentile range for a given region) and bottom-tier ZHVI (typical value for homes that fall within the 5th to 35th percentile range for a given region)."

ZHVI All Homes - Top tier time series $ (Metro and US)
```{r echo=FALSE}
data(zhvi_toptier)
kbl(head(zhvi_toptier[, 1:6]))
```

..

ZHVI All Homes - Bottom tier time series $ (Metro and US)
```{r echo=FALSE}
data(zhvi_bottomtier)
kbl(head(zhvi_bottomtier[, 1:6]))
```


## <ins> Public Safety </ins>

According to LA County Open Data initiative, Public Safety includes Sheriff, Fire, and other Emergency related datasets

## <ins> Complaints &/or Settlements </ins>

### National Police Funding Database
Source: https://policefundingdatabase.tminstituteldf.org/misconductsettlements

Data on:
- Federal Grants
- Police Consent Degrees
- Police Misconduct Complaints (Only Chicago, IL)
- Police Misconduct Settlemtns
- Policy changes via Policy


### Open Police
US wide data on police complaints
Home page: https://openpolice.org/
Source: https://github.com/flexyourrights/openpolice

I have no idea how to access the data. Still exploring


### NYPD Misconduct Complaint Database
Source: https://www.nyclu.org/en/campaigns/nypd-misconduct-database


### Citizens Police Data Project
Source: https://cpdp.co/
collects and publishes information about police misconduct in Chicago

### City of Tacoma
Source: https://data.cityoftacoma.org/Public-Safety/Complaint-Allegations-with-Findings/x46s-jtje

Source of complaint, "Citizen", no other information is given about who complaints. Other relevant information is the Finding of the complaint.

Types of allegation
```{r echo=FALSE}
data(complaint_cot)
kbl(unique(complaint_cot$Allegation))
```

..

Finding/ruling on complaint
```{r echo=FALSE}
data(complaint_cot)
kbl(unique(complaint_cot$Finding))
```


### CAP, Legal Aid Society
Source: https://www.capstat.nyc/commands/#10/40.7618/-73.5686


## <ins> Militarization </ins>


Militarization and Police Violence: The Case of the 1033 Program
Source: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/SDDXRT

Militarization Fails to Enhance Police Safety or Reduce Crime, But May Harm Police Reputationhttps://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VYPUVC

### ACLUM
Source: https://github.com/ACLUmass/ACLUM-Public-Datasets

Data from these states. Each working sheet has information.

California example
```{r message=FALSE, warning=FALSE, echo=FALSE}
data(dd1033_CA)
colnames(dd1033_CA)
```

Total acquisition sum USD 7583.282.

```{r}
kbl(head(dd1033_CA))
```


# Whats next?

 Civil Asset Forfeiture? - https://www.vera.org/publications/price-of-prisons-2015-state-spending-trends#prisonPopulationAndSpending20102015
 
 War on drugs - police action to drugs linked to profit/revenue for department: https://drugpolicy.org/issues/policing


