---
title: "open source datasets research"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{open source datasets research}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(USF)
```

The following is a descriptive document of the open source datasets that might be of use for the USF summer institute.

# <ins> On Police shootings/killings </ins>

### Washington Post Data - police shootings
Source https://github.com/washingtonpost/data-police-shootings

Raw data: `wpdata`
Filtered by states of test cities: `t_wpdata`

To filter by test cities:
```{r eval=FALSE, include=TRUE}
data(t_wpdata)
sf_wpdata <- t_wpdata %>% filter(city == "San Francisco")
ch_wpata <- t_wpdata %>% filter(city == "Chicago")
more_wpdata <- t_wpdata %>% filter(city == "Baltimore")
nyc_wpdata <- t_wpdata %>% filter(city == "New York")
dt_wpdata <- t_wpdata %>% filter(city == "Detroit")
kcmo_wpdata <- t_wpdata %>% filter(city == "Kansas City")
```


```{r echo=TRUE}
data(t_wpdata)
plot(t_wpdata$geometry)
```


### Fatal Encounters
Source https://fatalencounters.org/spreadsheets/

Raw data: `encounters`
Filtered by states of test cities: `t_encounters`

To filter by test cities:
```{r eval=FALSE, include=TRUE}
data(t_encounters)
sf_encounters <- t_encounters %>% filter(city_death == "San Francisco")
ch_encounters <- t_encounters %>% filter(city_death == "Chicago")
more_encounters <- t_encounters %>% filter(city_death == "Baltimore")
nyc_encounters <- t_encounters %>% filter(city_death == "New York")
dt_encounters <- t_encounters %>% filter(city_death == "Detroit")
kcmo_encounters <- t_encounters %>% filter(city_death == "Kansas City")
```

```{r echo=TRUE}
data(t_encounters)
plot(t_encounters$geometry)
```

### Mapping Police Violence
Source: https://mappingpoliceviolence.org/aboutthedata

Whole data set: `mpv`


- Worksheet 1 "2013-2020 Police Killings": `mpv_polkill`
Filter by states: `t_mpv_polkill`
Geocoded data including intersections (created with ArcGIS Pro 10)

To filter by test cities
```{r eval=FALSE, include=TRUE}
sf_mpv_polkill <- t_mpv_polkill %>% filter(City == "San Francisco")
ch_mpv_polkill <- t_mpv_polkill %>% filter(City == "Chicago")
more_mpv_polkill <- t_mpv_polkill %>% filter(City == "Baltimore")
nyc_mpv_polkill <- t_mpv_polkill %>% filter(City == "New York")
dt_mpv_polkill <- t_mpv_polkill %>% filter(City == "Detroit")
kcmo_mpv_polkill <- t_mpv_polkill %>% filter(City == "Kansas City")
```


```{r echo=TRUE}
data(t_mpv_polkill)
plot(t_mpv_polkill$geometry)
```


- Worksheet 2 "2013-2020 Killings by PD" : `mpv_polkillbypd`
Filter by states and city: `t_mpv_polkillbypd`
Chart by city (only for San Francisco, Baltimore, New York, and Detroit)

```{r echo=TRUE}
knitr::kable(head(t_mpv_killbypd %>% filter(City == "San Francisco")))
```


- Worksheet 3 "2013-2020 Killings by State": `mpv_polkillbypd`
Filter by states: `t_mpv_polkillbypd`
```{r echo=FALSE}
knitr::kable(t_mpv_polkillbypd)
```

- Worksheet 4 "Police Killings of Black Men"
```{r echo=FALSE}
data(mpv_polkilbm)
colnames(mpv_polkilbm)
```




### Million Dollar Hoods
Source: https://milliondollarhoods.pre.ss.ucla.edu/test-iframe-2/

LAPD 2012-2017 by Neighborhoods (Ranking)
```{r echo=FALSE}
data(mdh_lapd)
colnames(mdh_lapd)
```

LASD 2012-2017 by Neighborhoods (Ranking)
```{r echo=FALSE}
data(mdh_lasd)
colnames(mdh_lasd)
```


### Police 


# <ins> Policing in general </ins>


### Timeline
http://criticalresistance.org/policing-timeline/


### Spending

How much Us cities spend on policing
https://www.statista.com/chart/10593/how-much-do-us-cities-spend-on-policing/

PopularDemocracy.org
https://www.populardemocracy.org/news-and-publications/congress-must-divest-billion-dollar-police-budget-and-invest-public-education


### Stops - Open Policing Stanford
Source: https://openpolicing.stanford.edu/data/
CSV and shapefiles for select locations.


Los Angeles example
```{r echo=FALSE}
data(laops)
colnames(laops)
```


### Police in schools


Policing America’s Schools: An Education Week Analysis
U.S. and States
Source: https://www.edweek.org/which-students-are-arrested-most-in-school-u-s-data-by-school#/overview

### Jails and Prisons


### National Neighbourhood Watch
https://nnw.org/find-a-watch-program?type=2&field_county_value=&field_address_locality=&field_address_postal_code=&field_address_administrative_area=


### Critical Resistance - Abolishing Policing
Source: http://criticalresistance.org/abolish-policing/

#### On the Road with Abolition - ppt https://docs.google.com/presentation/d/1Vs1Dt53DR6s2vShPZw5xNX-QCDC2Ppws07ctl6h_Weg/edit#slide=id.g80c2475b0e_1_80

"the most basic police interaction, the stop"


# <ins> Drawing the Black Box </ins>

## <ins> Property data </ins>

Most cities have property data available you can download for free: "Property Vaulation and Assessment Data". 

### NYC Real Estate Assessment Property data.
Built with the Department of Finance values.
Source: https://data.cityofnewyork.us/City-Government/Property-Valuation-and-Assessment-Data/yjxr-fw8i

### MassGIS - MA Interactive Property Map
Datasets per city ready for download.
Source: https://massgis.maps.arcgis.com/apps/OnePane/basicviewer/index.html?appid=47689963e7bb4007961676ad9fc56ae9

### LA County Open Data Initiative
Property related data: https://data.lacounty.gov/browse?category=Property%2FPlanning&utf8=%E2%9C%93

### Zillow
Source: https://www.zillow.com/research/data/

"Zillow publishes top-tier ZHVI (typical value for homes within the 65th to 95th percentile range for a given region) and bottom-tier ZHVI (typical value for homes that fall within the 5th to 35th percentile range for a given region)."

ZHVI All Homes - Top tier time series $ (Metro and US)
```{r echo=FALSE}
data(zhvi_toptier)
kbl(head(zhvi_toptier[, 1:6]))
```

..

ZHVI All Homes - Bottom tier time series $ (Metro and US)
```{r echo=FALSE}
data(zhvi_bottomtier)
kbl(head(zhvi_bottomtier[, 1:6]))
```


## <ins> Public Safety </ins>

According to LA County Open Data initiative, Public Safety includes Sheriff, Fire, and other Emergency related datasets

## <ins> Complaints &/or Settlements </ins>

### National Police Funding Database
Source: https://policefundingdatabase.tminstituteldf.org/misconductsettlements

Data on:
- Federal Grants
- Police Consent Degrees
- Police Misconduct Complaints (Only Chicago, IL)
- Police Misconduct Settlemtns
- Policy changes via Policy

Example of an incomplete database.

### New York State Nuisance Abatement Actions
Source: https://www.propublica.org/datastore/dataset/new-york-state-supreme-court-nuisance-abatement-actions

Data is $200.


### Open Police
US wide data on police complaints
Home page: https://openpolice.org/
Source: https://github.com/flexyourrights/openpolice

I have no idea how to access the data. Still exploring


### NYPD Misconduct Complaint Database
Source: https://www.nyclu.org/en/campaigns/nypd-misconduct-database


### Citizens Police Data Project
Source: https://cpdp.co/
collects and publishes information about police misconduct in Chicago

### City of Tacoma
Source: https://data.cityoftacoma.org/Public-Safety/Complaint-Allegations-with-Findings/x46s-jtje

Source of complaint, "Citizen", no other information is given about who complaints. Other relevant information is the Finding of the complaint.

Types of allegation
```{r echo=FALSE}
data(complaint_cot)
kbl(unique(complaint_cot$Allegation))
```

..

Finding/ruling on complaint
```{r echo=FALSE}
data(complaint_cot)
kbl(unique(complaint_cot$Finding))
```


### CAP, Legal Aid Society
Source: https://www.capstat.nyc/commands/#10/40.7618/-73.5686


## <ins> Militarization </ins>


Militarization and Police Violence: The Case of the 1033 Program
Source: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/SDDXRT

Militarization Fails to Enhance Police Safety or Reduce Crime, But May Harm Police Reputationhttps://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VYPUVC

### ACLUM
Source: https://github.com/ACLUmass/ACLUM-Public-Datasets

Data from these states. Each working sheet has information.

California example
```{r message=FALSE, warning=FALSE, echo=FALSE}
data(dd1033_CA)
colnames(dd1033_CA)
```

Total acquisition sum USD 7583.282.

```{r}
kbl(head(dd1033_CA))
```


# Whats next?

 Civil Asset Forfeiture? - https://www.vera.org/publications/price-of-prisons-2015-state-spending-trends#prisonPopulationAndSpending20102015
 
 War on drugs - police action to drugs linked to profit/revenue for department: https://drugpolicy.org/issues/policing


